---
title: 'MECH481A6: Engineering Data Analysis in R'
subtitle: 'Chapter 10 Homework: Measurement' 
author: 'Student Name'
date: '`r format(Sys.time(), "%d %B, %Y")`'
output: pdf_document
---

```{r global-options, include=FALSE}
# set global options for figures, code, warnings, and messages
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.path="../figs/",
                      echo=FALSE, warning=FALSE, message=FALSE)
```

# Load packages

```{r load-packages, message=FALSE}
# load packages for current session
library(tidyverse) 
library(gridExtra) 
library(MASS)
library(lubridate)
```

# Chapter 10 Homework

This homework will give you practice at working with a measurement dataset: `airlift_mass_repeatability.csv`. This data set represents repeated measures of "blank" air sampling filters. 

A couple notes to consider when reporting answers in response to questions. The microbalance used to make these measurements reads out to the nearest microgram ($\mu g$), which is 0.000001 $g$ or 0.001 $mg$. Thus, be careful when reporting descriptive statistics so as not to overstate your **precision**. Use the `round()` function to avoid reporting more than 0.1 $\mu g$ of precision (or 0.0001 $mg$). Here is some example code that uses the `across()` function from `dplyr::` to round numeric output to just four digits (appropriate for $mg$ units in this exercise):

`dplyr::mutate(across(.cols = where(is.numeric), .fns = round, 3))`

\newpage

## Question 1
Import the `airlift_mass_repeatability.csv` file into a data frame called `blanks` and perform the following data wrangling in a single pipe:  

- retain only the first 3 columns of data;
- rename the columns with the names `date`, `id`, and `mass_mg`;
- convert the `date` column vector into a date class object using `lubridate::`
- convert the `id` variable to a class `factor` (this can be accomplished using `base::as.factor()` or `purrr::as_factor()`)
- create a new column vector named `mass_mg` by rescaling the `mass_g` data (i.e., convert $g$ to $mg$ by multiplying `mass_g` by 1000)

```{r import-clean-data}
blanks <- read_csv("/Users/Maxwell Grisham/Documents/Grisham-MECH476/Homework/AIRLIFT_mass_repeatability.csv") %>%
  dplyr::select(1:3) %>%
  rename(date = 1, id = 2, mass_g = 3) %>%
  mutate(date = dmy(date)) %>%
  mutate(id = as.factor(id)) %>%
  mutate(mass_mg = mass_g * 1000)

```

## Question 2:  

  2a. Are there any NAs present in the data frame?  
  -There are zero NA values in the data frame
  2b. How many unique filter IDs are present in this data frame?  
  -There are five unique filter IDs
  2c. How many samples are present for each filter ID? Hint: look up the `dplyr::count()` function.
  -41666 = 78
  41667 = 78
  41668 = 78
  41669 = 76
  41671 = 78
  2d. Over how long of a period were these blank measurements made? Hint: this can be done in base R with a `max() - min()` or with `lubridate::interval() %>% as.duration()`.
  -5 Weeks
```{r unique-IDs}
num_NA <- colSums(is.na(blanks))

id_unique <- unique(blanks$id)

id_sample <- blanks %>%
  count(id)

period <- interval(min(blanks$date), max(blanks$date)) %>%
  as.duration()
period
```

## Question 3

Group the `blanks` data frame by `id` and calculate mean, median, and standard deviations for each filter id.  Hint: use `group_by() %>% summarise()` to do this efficiently.

```{r blank-descriprives}
blanks_summary <- blanks %>%
  group_by(id) %>%
  summarise(
    mean_mass = mean(mass_mg),  #XL: please pay attention to the  **precision**. Use the `round()` function to avoid reporting more than 0.1 $\mu g$ of precision (or 0.0001 $mg$). For example, round(mean(mass_mg), digits = 3). Same for Question 3.
    median_mass = median(mass_mg),  
    sd_mass = sd(mass_mg)
  )
blanks_summary

```

## Question 4

Calculate the limit of detection (LOD) for this measurement method. Note: you will need to calculate standard deviations for each filter `id` (as done in question 3) and then estimate LOD from $LOD = 3\cdot \sigma_b$ where $\sigma_b$ is calculated for each filter `id`.

```{r LOD}
blanks_summary <- blanks %>%
  group_by(id) %>%
  summarise(
    mean_mass = mean(mass_mg),
    median_mass = median(mass_mg),
    sd_mass = sd(mass_mg)
  )
blanks_summary

blanks_LOD <- blanks %>%
  group_by(id) %>%
  summarise(
    sd_mass = sd(mass_mg),
    LOD = 3 * sd_mass
  )

blanks_LOD
```
