---
title: 'MECH476: Engineering Data Analysis in R'
subtitle: 'Chapter 7 Homework: Multivariate Exploratory Data Analysis' 
author: 'Student Name'
date: '`r format(Sys.time(), "%d %B, %Y")`'
output: pdf_document
---

```{r global-options, include=FALSE}
# set global options for figures, code, warnings, and messages
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.path="../figs/",
                      echo=FALSE, warning=FALSE, message=FALSE)
```

# Load packages

```{r load-packages, message=FALSE}
# load packages for current session
library(tidyverse)
library(readr)
library(ggplot2)
library(janitor)
```

# Chapter 7 Homework

In [Chapter 5](https://smogdr.github.io/edar_coursebook/eda1.html#cumulative-distribution-plot), we briefly explored data on the salaries of engineering
graduates from the National Science Foundation 2017 [National Survey of College Graduates](https://www.nsf.gov/statistics/srvygrads/) from a univariate
perspective. Now, let's explore the relationships between multiple variables. 

When a question asks you to make a plot, remember to set a theme, title, 
subtitle, labels, colors, etc. It is up to you how to personalize your plots, 
but put in some effort, and make the plotting approach consistent throughout
the document. For example, you could use the same theme for all plots. I also
like to use the subtitle as a place for the main summary for the viewer.

\newpage

## Question 1: Data wrangling

Within a pipeline, import the data from the .csv file, convert all column names
to lowercase text (either "manually" with `dplyr::rename()`, or use
[`clean_names()` from the `janitor` package](https://rdrr.io/cran/janitor/man/clean_names.html)), convert `gender`
from "numeric" to "factor", and drop any and all observations with `salary` 
recorded as 0. Assign this to a dataframe object with a meaningful name. 

```{r import-data}
# import and tidy salary data
data_raw <- read_csv("/Users/Maxwell Grisham/Documents/Grisham-MECH476/ME_salaries.csv") 
data_clean <- data_raw %>%
  clean_names() %>%
  filter(salary > 0)
data_clean$gender <- factor(data_clean$gender)  #XL: You can use pipe and mutate to combine this line of code with your former code. %>% mutate(gender = factor(gender))
```

How many observations have a 0 (zero) value for salary? Note: The last question asked you to remove these observations from the resultant data frame.
-15 observations with a zero value for salary.

```{r salary-zero}
# number of observations with salary as 0 
zero_salary <- nrow(data_raw) - nrow(data_clean) #XL: Is there any NAs in your data_raw? If yes, this calculation will include NAs in your result. You can use filter(salary == 0) %>% nrow().
zero_salary
```

What are the levels in `gender`? (Ignore the fact that the observations
refer to "biological sex", not "gender". *Gender* is now recognized as a fluid
term with more than two options; *biological sex* - what was assigned at birth - 
is binary term).
-There are two levels "F" and "M"
```{r gender-levels}
# number of factor levels
levels(data_clean$gender)
```

## Question 2: Univariate EDA

Using what you learned in Chapter 5, generate basic plots and/or descriptive 
statistics to explore `age`, `gender`, and `salary`. List whether each variable
is continuous or categorical, and explain how and why you adjusted your EDA
approach accordingly.

-I have made comparisons focusing on salary and age with the gender being the fill, 
salary and age are continuous variables while gender is a categorical variable.

```{r univariate-eda, include=TRUE}
# univariate eda
data_plot <- data_raw %>%
  clean_names() %>%
  filter(salary > 0)

plot1 <- ggplot(data = data_plot,
       aes(x = age,
       fill = gender)) +
  geom_histogram() +
  theme_minimal()
plot1

plot2 <- ggplot(data = data_plot,
                aes(x = age,
                    y = salary,
                    color = gender)) +
  geom_point(alpha = 0.5) +
  theme_minimal()

plot2

plot3 <- ggplot(data = data_plot,
                aes(x = salary,
                    fill = gender)) + 
  geom_histogram(binwidth = 10000) +
  theme_minimal()
plot3

plot4 <- ggplot(data = data_plot,
                aes(y = gender,
                    x = salary,
                    fill = gender)) + 
  geom_boxplot()
```

\newpage

## Question 3: Multivariate histograms

Create a histogram of `salary`, faceted by `gender`. Add `bins = 50` and 
`color = "lightgrey"`.

```{r hist-salary}
# histogram of salaries split by gender
ggplot(data = data_plot,
                aes(x = salary,
                    fill = gender)) + #XL: Please try to use facet, e.g., facet_wrap(~gender), even though fill = gender has the same function here.
  geom_histogram(binwidth = 10000) +
  theme_minimal()
```

Create a histogram of `age`, faceted by `gender`. Add `bins = 50` and 
`color = "lightgrey"`.

```{r hist-age}
# histogram of ages split by gender
ggplot(data = data_plot,
       aes(x = age,
       fill = gender)) +     #XL: Please try to use facet, e.g., facet_wrap(~gender), even though fill = gender has the same function here.
  geom_histogram() +
  theme_minimal()
```

\newpage

## Question 4: Multivariate boxplots

Create a boxplot of `salary`, faceted by `gender`. Use `oulier.shope = 1` to 
better visualize the outliers.

```{r salary-box}
# boxplots of salary data by gender
ggplot(data = data_plot,
                aes(x = gender,
                    y = salary,
                    fill = gender)) + 
  geom_boxplot(outlier.shape = 1) +
  facet_wrap(~ gender)
  theme_minimal()
```

Create a boxplot of `age`, faceted by `gender`.

```{r age-box}
# boxplots of age data by gender
ggplot(data = data_plot,
                aes(x = gender,
                    y = age,
                    fill = gender)) + 
  geom_boxplot(outlier.shape = 1) +
  facet_wrap(~ gender) +
  theme_minimal()
```

\newpage

## Question 5: Scatterplot and correlation

Create a scatterplot of `age` (x-axis) and `salary`, differentiating by 
`gender`. 

```{r multivar-scatter}
# scatterplot of salary across age by gender
ggplot(data = data_plot,
       aes(x = age,
           y = salary,
           color = gender)) +     #XL: Many points are overlaped on this plot, so it would be nice to use facet to saperate this plot into two panels by gender.
  geom_point(alpha = 0.5) +
  theme_minimal()
```

*Bonus point*: Is there a correlation between an engineer's salary and age?
What is the estimated Pearson correlation coefficient *r*? Run a formal test.

```{r corr-test}
# correlation test
#XL: data_plot %>%
#XL:   group_by(gender) %>%
#XL:   summarise(correlation = cor(age, salary))
```

\newpage

## Question 6: Cumulative distribution function

Plot the cumulative distribution function of `salary` by `gender`. Adjust the
x-axis with `scale_x_log10(limits = c(5e4, 5e5))` to zoom in a bit. What do you
notice about the salaries for men and women? Hint: Remember there are greater
differences the farther up you go on a log scale axis.

```{r salary-ecdf}
# plot cdf of salary by gender
ggplot(data = data_plot, aes(x = salary, color = gender)) +
  stat_ecdf(geom = "step") +
  scale_x_log10(limits = c(5e4, 5e5)) +
  theme_minimal() +
  labs(x = "Salary (log scale)", y = "Cumulative Probability", 
       title = "CDF of Salary by Gender")
```

## Question 7: Quantiles

Calculate the quantiles of `salary` by `gender`. You can either subset the data
with `dplyr::filter()` and dataframe assignment, or you can group by, summarize
by quantile, and ungroup. 

*Bonus point*: Assign the output to a dataframe, and use inline code to call
individual values when answering the following questions. Do not let R use
scientific notation in the text output; check the knitted document.

```{r salary-quantiles, message=FALSE}
# calculate quantiles of salary by gender
salary_male <- data_clean %>%
  filter(gender == "M")
salary_female <- data_clean %>%
  filter(gender == "F")

male_quant <- quantile(salary_male$salary, probs = seq(0, 1, by = 0.25))
female_quant <- quantile(salary_female$salary, probs = seq(0, 1, by = 0.25))

print(male_quant)
print(female_quant)

```

What is the difference in salary between men and women at the median? 
-difference is $7000

  * Median salary for women is
  -$90000
  * Median salary for men is 
  -$97000
  * The difference at the median is 
  -$7000
  
At the top percentile (maximum)? 

  * Maximum salary for women is 
  -$350000
  * Maximum salary for men is 
  -1027653
  * The difference at the maximum is 
  -$677653

Do you think there is a salary difference by gender across the pay scale? What
other information would you need to test your hypothesis?

-I do believe there is a salary difference across the pay scale, the other information 
I would need is a more equal division between males and females, as well as take
into account how long they have been working in the field.
## Question 8: Hypothetical analysis

Think about what other variables you would like to include in an hypothetical
analysis. From your perspective, what are the most important individual,
family, and workforce factors related to salary---beyond gender and age?
-Amount of experience, race/ethnicity, education level, and location.
\newpage

## Question 9: Recreate plot

Recreate this plot with the `mpg` dataset. Remember to use `?mpg` for 
information on the dataset and the variables. How would you describe the 
correlation between the independent variable and dependent variable? Do you
see any patterns when considering the third variable?

(View R Markdown PDF for image)

```{r mpg-plot}
# call mpg pdf - you need to recreate it 
ggplot(data = mpg, aes(x = displ, y = hwy, color = class)) +
  stat_ecdf(geom = "smooth")
  #geom_smooth(se = FALSE, aes(group = class), method = "loess") +  # Add smoothed lines without confidence intervals
  #geom_point() +  # Add points for better visibility
  theme_minimal() 

knitr::include_graphics("./mpg-ch7-plot.pdf")
```

\newpage

# Appendix

```{r show_code, ref.label = knitr::all_labels(), echo = TRUE, eval = FALSE}
```
